Community engagement and subgroup meta-knowledge: Some factors in the soul of a community

### 2018-12-28 PaperFinalDraft.Rnw
update affiliation
Many changes, including: -adding citations for R packages -adding paragraph about reproducibility -fixing non-square tiles in correlation plot -adjusting affiliation back to grad school -fixing axis labels for Figure 5

\author{Amelia McNamara}
\institute{A. McNamara \at
              Burton 215 \\
              Smith College
              Northampton, MA 01063 \\
              \email{amcnamara@smith.edu}
}

@@

This article is one of several related to the Knight Foundation community data from the 2013 American Statistical Association Data Expo. For more information on the Expo and the data sets, see \citep{Hof2019}. This article is fully-reproducble, with all code necessary to generate the paper available on GitHub~\citep{McN2019}. In the GitHub repository, you will find the underlying data, the LaTeX/knitr file which produces the paper including R code, a packrat archive of the necessary R packages, and LaTeX style files to create the specific journal style.

@@

<<YNall, out.width="0.99\\linewidth",fig.width = 15,fig.height=9.5,fig.cap="Responses to yes/no questions about participants' behaviors, comparing all three survey years.">>=
baseplot <- ggplot(mapping = aes(Question, Freq, fill = Response)) + 
  facet_wrap(~Year) + 
  geom_bar(data = trial1$neg, stat = "identity") + 
  geom_bar(data = trial1$pos, stat = "identity") + 
  scale_y_continuous(
    breaks=seq(from=-0.75, to=0.75, by=0.25), 
    labels=percent
    ) + 
  coord_flip()+ 
  xlab("") + 
  scale_x_discrete(
    labels=rev(c("Registered to vote \n", "Voted in local election \n", "Donated money to a \n local organization", "Attended a local \n festival or event", "Gave money or food \n to an individual", "Participated in a \n church event", "Performed local \n volunteer work", "Worked with other residents \n to make change", "Attended a local \n public meeting", "Provided free shelter \n to an individual"))
    ) + 
  ylab("") + 
  theme(
    legend.text=element_text(size=14), 
    legend.title=element_text(size=16), 
    axis.text.y=element_text(size=20), 
    title=element_text(size=24), 
    axis.text.x=element_text(size=14), 
    strip.text=element_text(size=18))
baseplot
@

@@

<<YN2010plot, out.width="0.99\\linewidth",fig.width = 10, fig.height=18, fig.cap="Percentage difference from overall survey rates (2010 data). The letters A-J represent the activities listed in Figure \\ref{fig:YNall}. A: Registered to vote, B: Voted in a local election, C: Donated money to a local organization, D: Attended a local event, E: Gave money or food to an individual, F: Participated in a church event, G: Performed local volunteer work, H: Worked with other residents to make change, I: Attended a local public meeting, J: Provided free shelter to an individual.">>=
baseplot <- ggplot(dfcitiesYes) + 
  aes(Question, diff, fill = col) + 
  facet_wrap(~citystate, ncol=3) + 
  geom_bar(stat = "identity") + 
  coord_flip() + 
  guides(fill=FALSE) + 
  scale_y_continuous(
    breaks=seq(from=-.5, to=.5, by=.125),
    labels=c("50%","", "25%","", "0","", "25%","", "50%")
    ) +
  scale_x_discrete(labels=LETTERS[10:1]) + 
  ylab("") + 
  xlab("") + 
  theme(
    axis.text.y=element_text(size=12), 
    title=element_text(size=18), 
    axis.text.x=element_text(size=14), 
    strip.text=element_text(size=16))
baseplot
@

@@

cp <- ggplot(data=melted, aes(x=Var1, y=Var2, fill=value)) +
  geom_tile() + 
  scale_fill_gradient2(low="#762A83", high="#1B7837", name="Correlation") + 
  scale_x_discrete(labels=varnames) + 
  scale_y_discrete(labels=varnames) + 
  xlab("") + 
  ylab("") + 
  theme(axis.text.x=element_text(angle=-90, hjust=0)) + 
  coord_flip()
cp
@

@@

\section{Conclusions and Further Work}
Through this exploration of the Knight Foundation's Soul of the Community data, we have been able to examine variability between communities and theorize about what might affect those variations. In Section \ref{communittsatsec} we saw which communities reported the highest levels of community satisfaction, namely State College, PA and Boulder, CO. In Section \ref{behaviorsec}, we explored the common behaviors that survey participants engaged in, including registering to vote and voting in local elections. We noticed that college towns tend to have higher than typical rates of providing shelter to a non-relative (the couchsurfing effect?).

>>>>

\author{Amelia A McNamara}
\institute{A. McNamara \at
              University of California, Los Angeles \\
              Los Angeles, CA 90095 \\
              \email{amelia.mcnamara@stthomas.edu}
}

@@

This article is one of several related to the Knight Foundation community data from the 2013 American Statistical Association Data Expo. For more information on the Expo and the data sets, see \citep{Hof2019}. This article is fully reproducble. The paper was written using \LaTeX\, and R~\citep{R2014}. The R package knitr was used to interweve writing and R code~\citep{Xie2014}. If you inspect the source code, available on Github~\citep{McN2019}, you will find that each of the figures in the paper are generated when the document is compiled (``knit''). Likewise, any statistic mentioned in the text is calculated using R and automatically inserted into the text using knitr.  All graphics were generated with the package ggplot2, an R implementation of The Grammar of Graphics~\citep{Wic2009, Wil2005}. The paper was written in the time period between the development of the R package plyr~\citep{Wic2011} and the R package dplyr, which superceded it~\citep{WicFra2015}. Therefore, it uses functions from both packages. One challenge with reproducibility is that as packages develop, they are not always backwards-compatible. To help combat this issue, the code available on Github includes a packrat library. The packrat package allows a user to snapshot the state of their packages in order to freeze them for future reproducibility~\citep{UshMcP2014}. For example, although the current version of dplyr is 0.7.8, this paper uses version 0.4.1. In order to run the code hosted on Github, you will need to ``hydrate'' the library of packages using packrat. The Github repository contains the underlying data, the LaTeX/knitr file which produces the paper (including R code), a packrat archive of the necessary R packages, and LaTeX style files to create the specific journal style.

The remainder of this paper is an exploration of the idea of community meta-knowledge. It begins with discussion of the data in Section \ref{data}, then moves to focus on community satisfaction in Section \ref{communittsatsec}, and community engagement through local behaviors in Section \ref{behaviorsec}. Finally, it discusses meta-knowledge in Section \ref{metasec} and wraps up with conclusions in Section \ref{conclusion}.

@@

<<YNall, out.width="0.99\\linewidth",fig.width = 15,fig.height=9.5,fig.cap="Responses to yes/no questions about participants' behaviors, comparing all three survey years.">>=
baseplot <- ggplot(mapping = aes(Question, Freq, fill = Response)) + 
  facet_wrap(~Year) + 
  geom_bar(data = trial1$neg, stat = "identity") + 
  geom_bar(data = trial1$pos, stat = "identity") + 
  scale_y_continuous(
    breaks=seq(from=-0.75, to=0.75, length.out = 7), 
    labels=c("", "-50%", "", "0","",  "50%", "")
    ) + 
  coord_flip()+ 
  xlab("") + 
  scale_x_discrete(
    labels=rev(c("Registered to vote \n", "Voted in local election \n", "Donated money to a \n local organization", "Attended a local \n festival or event", "Gave money or food \n to an individual", "Participated in a \n church event", "Performed local \n volunteer work", "Worked with other residents \n to make change", "Attended a local \n public meeting", "Provided free shelter \n to an individual"))
    ) + 
  ylab("") + 
  theme(
    legend.text=element_text(size=14), 
    legend.title=element_text(size=16), 
    axis.text.y=element_text(size=20), 
    title=element_text(size=24), 
    axis.text.x=element_text(size=14), 
    strip.text=element_text(size=18))
baseplot
@

@@

<<YN2010plot, out.width="0.99\\linewidth",fig.width = 10, fig.height=18, fig.cap="Difference from overall survey rates (2010 data). The letters A-J represent the activities listed in Figure \\ref{fig:YNall}. A: Registered to vote, B: Voted in a local election, C: Donated money to a local organization, D: Attended a local event, E: Gave money or food to an individual, F: Participated in a church event, G: Performed local volunteer work, H: Worked with other residents to make change, I: Attended a local public meeting, J: Provided free shelter to an individual.">>=
baseplot <- ggplot(dfcitiesYes) + 
  aes(Question, diff, fill = col) + 
  facet_wrap(~citystate, ncol=3) + 
  geom_bar(stat = "identity") + 
  coord_flip() + 
  guides(fill=FALSE) + 
  scale_y_continuous(
    breaks=seq(from=-.15, to=.15, length.out = 5),
    labels=c("", "-7.5%", "0","+7.5%",  "")
    ) +
  scale_x_discrete(labels=LETTERS[10:1]) + 
  ylab("") + 
  xlab("") + 
  theme(
    axis.text.y=element_text(size=12), 
    title=element_text(size=18), 
    axis.text.x=element_text(size=11), 
    strip.text=element_text(size=16))
baseplot
@

@@

cp <- ggplot(data=melted, aes(x=Var1, y=Var2, fill=value)) +
  geom_tile() + 
  scale_fill_gradient2(low="#762A83", high="#1B7837", name="Correlation") + 
  scale_x_discrete(labels=varnames) + 
  scale_y_discrete(labels=varnames) + 
  xlab("") + 
  ylab("") + 
  theme(axis.text.x=element_text(angle=-90, hjust=0)) + 
  coord_flip() +
  coord_fixed()
cp
@

@@

\section{Conclusions and Further Work}
\label{conclusion}
Through this exploration of the Knight Foundation's Soul of the Community data, we have been able to examine variability between communities and theorize about what might affect those variations. In Section \ref{communittsatsec} we saw which communities reported the highest levels of community satisfaction, namely State College, PA and Boulder, CO. In Section \ref{behaviorsec}, we explored the common behaviors that survey participants engaged in, including registering to vote and voting in local elections. We noticed that college towns tend to have higher than typical rates of providing shelter to a non-relative (the couchsurfing effect?).
